Examining background gene lists in clusterProfiler - RNA-seq example

### 2024-05-31 analysis/analyse_rnaseq1.Rmd
fix
better gene set stats

lapply(gs,length)

lapply(gs,function(x) {
  y <- unlist(lapply(x, length ))
  summary(y)
} )

lapply(gs,function(x) {
  y <- unique(unlist(x))
  length(y)
} )

@@

barplot(sort(jac2),horiz=TRUE,las=1,xlim=c(0,1),
  main="Jaccard similarity Corrected:Original",xlab="fold increase")

>>>>

# number of sets
lapply(gs,length)

# set size summary
lapply(gs,function(x) {
  y <- unlist(lapply(x, length ))
  summary(y)
} )

# gene coverage
lapply(gs,function(x) {
  y <- unique(unlist(x))
  length(y)
} )

# total number of annotations
lapply(gs, function(x) {
  sum(unlist(lapply(x,length)))
} )

@@

barplot(sort(jac2),horiz=TRUE,las=1,xlim=c(0,1),
  main="Jaccard similarity Corrected:Original",xlab="Jaccard index")

### 2024-05-31 analysis/analyse_rnaseq2.Rmd
fix
better gene set stats

  gs1 <- gs1[which(gs1$gene %in% bg),]

@@

# bad=ORA analysis, a list with results of up and down regulated pathways
# good=same as bad, except fixed the background problem
compare_ora_plots <- function(bad,good){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  mup <- merge(badup,goodup,by=0)
  MAX=max(c(mup$es.x,mup$es.y))

  plot(mup$es.x, mup$es.y,
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original", ylab="corrected", main="Enrichment score")
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  MAX=max(-log10(c(mup$pvalue.x,mup$pvalue.y)))

  plot(-log10(mup$pvalue.x), -log10(mup$pvalue.y),
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original", ylab="corrected", main="-log10 p-value")
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  mdn <- merge(baddn,gooddn,by=0)
  MAX=max(c(mdn$es.x,mdn$es.y))

  plot(mdn$es.x, mdn$es.y,
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original", ylab="corrected", main="Enrichment score")
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  MAX=max(-log10(c(mdn$pvalue.x,mdn$pvalue.y)))

  plot(-log10(mdn$pvalue.x), -log10(mdn$pvalue.y),
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original", ylab="corrected", main="-log10 p-value")
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)
}

compare_ora <- function(bad,good){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  mup <- merge(badup,goodup,by=0)
  mdn <- merge(baddn,gooddn,by=0)
  bup <- subset(mup,p.adjust.x<0.05)$Row.names
  gup <- subset(mup,p.adjust.y<0.05)$Row.names
  bdn <- subset(mdn,p.adjust.x<0.05)$Row.names
  gdn <- subset(mdn,p.adjust.y<0.05)$Row.names
  sig_bad <- unique(c(bup,bdn))
  sig_good <- unique(c(gup,gdn))
  nsig_bad <- length(sig_bad)
  nsig_good <- length(sig_good)
  jac <- length(intersect(sig_bad,sig_good)) / length(union(sig_bad,sig_good))
  result=c("nsig_bad"=nsig_bad,"nsig_good"=nsig_good,"jaccard"=jac)
  return(result)
}

@@

ora_dat <- lapply(1:3, function(i) {
  j=params[i,1]
  k=params[i,2]
  bad <- orabad(d[[j]],gs[[k]])
  good <- oragood(d[[j]],gs[[k]])
  res <- compare_ora(bad,good)
  return(res)
} )

@@

barplot(sort(jac2),horiz=TRUE,las=1,xlim=c(0,1),
  main="Jaccard similarity Corrected:Original",xlab="fold increase")

>>>>

@@

# bad=ORA analysis, a list with results of up and down regulated pathways
# good=same as bad, except fixed the background problem
compare_ora_plots <- function(bad,good){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  mup <- merge(badup,goodup,by=0)

  MAX=max(-log10(c(mup$p.adjust.x,mup$p.adjust.y)))

  plot(-log10(mup$p.adjust.x), -log10(mup$p.adjust.y),
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original", ylab="corrected", main="-log10 FDR")
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  mdn <- merge(baddn,gooddn,by=0)

  MAX=max(-log10(c(mdn$p.adjust.x,mdn$p.adjust.y)))

  plot(-log10(mdn$p.adjust.x), -log10(mdn$p.adjust.y),
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original", ylab="corrected", main="-log10 FDR")
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)
}

compare_ora <- function(bad,good){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  bup <- rownames(subset(badup,p.adjust<0.05))
  gup <- rownames(subset(goodup,p.adjust<0.05))
  bdn <- rownames(subset(baddn,p.adjust<0.05))
  gdn <- rownames(subset(gooddn,p.adjust<0.05))
  sig_bad <- unique(c(bup,bdn))
  sig_good <- unique(c(gup,gdn))
  nsig_bad <- length(sig_bad)
  nsig_good <- length(sig_good)
  jac <- length(intersect(sig_bad,sig_good)) / length(union(sig_bad,sig_good))
  result=c("nsig_bad"=nsig_bad,"nsig_good"=nsig_good,"jaccard"=jac)
  return(result)
}

@@

@@

barplot(sort(jac2),horiz=TRUE,las=1,xlim=c(0,1),
  main="Jaccard similarity Corrected:Original",xlab="Jaccard index")

# overall
oa <- matrix(paste(res1$nsig_bad,res1$nsig_good,signif(res1$jaccard,2)),ncol=9)
rownames(oa) <- names(d)
colnames(oa) <- names(gs)

oa %>%
  kbl(row.names = TRUE, caption="Original:Corrected:Jaccard") %>%
  kable_paper("hover", full_width = F)
